%================================================================================
%       Safety Critical Systems Club - Data Safety Initiative Working Group
%================================================================================
%                       DDDD    SSSS  IIIII  W   W   GGGG
%                       D   D  S        I    W   W  G   
%                       D   D   SSS     I    W W W  G  GG
%                       D   D      S    I    WW WW  G   G
%                       DDDD   SSSS   IIIII  W   W   GGG
%================================================================================
%               Data Safety Guidance Document - LaTeX Source File
%================================================================================
%
% Description:
%   Acknowledgements section.
%
%================================================================================
\section{Acknowledgements (Discursive)} \label{bkm:acknowledgements}

\dsiwgSectionQuote{Our ability to do great things with data will make a real difference in every aspect of our lives.}{Jennifer Pahlka}

The document contributors would like to thank:
\begin{itemize}
  \item The \glsfirst{scsc}.
  \item The \gls{scsc} Covid-19 Working Group for providing some of the data used in the Covid-19 Appendix.
  \item Brian Jepson of the \gls{scsc} for web hosting support and technical help with the \gls{scsc} web site.
  \item
    Mark Templeton for editing this edition.
  \item Paul Hampton
    and
    Mark Templeton for managing the publication processes.
  \item
    \cbstart
    Nick Hales, Mike Parsons, Tim Rowe, Alan Simpson and Mark Templeton
    for developing the additional text for this edition.
    \cbend
  \item
    Martin Atkins and Divya Atkins for driving the development of tooling and promoting data safety.
  \item Mike Parsons for chairing the Working Group meetings.
  \item All those who have taken minutes at Working Group meetings.
  \item
    All the organisations that have hosted Working Group meetings.
  \item All the organisations that have provided support to the document's contributors.
  \item Those that have been unable to attend meetings but have made supporting contributions.
\end{itemize}
