%================================================================================
%       Safety Critical Systems Club - Data Safety Initiative Working Group
%================================================================================
%                       DDDD    SSSS  IIIII  W   W   GGGG
%                       D   D  S        I    W   W  G   
%                       D   D   SSS     I    W W W  G  GG
%                       D   D      S    I    WW WW  G   G
%                       DDDD   SSSS   IIIII  W   W   GGG
%================================================================================
%               Data Safety Guidance Document - LaTeX Source File
%================================================================================
%
% Description:
%   Definitions and Abbreviations (Normative) section.
%
%================================================================================
\section{Definitions (Normative)} \label{bkm:definitionsabbreviations}

\dsiwgSectionQuote{I love data. I think it's very important to get it right, and I think it's good to question it.}{Mary Meeker}

\printnoidxglossary[type=definitions,style=altlist,nonumberlist]

\subsection{Definitions}
\begin{longtable}{|L{\dsiwgColumnWidth{0.35}}|L{\dsiwgColumnWidth{0.65}}|}
  \caption{\protect Normative definitions\protect}
  \\\hline\TableHeadColour{Term} & \TableHeadColour{Definition}\\\hline
  \endfirsthead
  \caption[]{Basic definitions (continued)}
  \\\hline\TableHeadColour{Term} & \TableHeadColour{Definition}\\\hline
  \endhead
  \multicolumn{2}{r}{\sl Continued on next page}
  \endfoot
  \endlastfoot
	{Data Artefact\index{Artefact, Data|textbf}} & {An item, or collection of items, that provides a useful perspective on data generated, processed or consumed by a system.}\\\hline
	{\index{Data!Owner|textbf}Data Owner} & {The individual or organization responsible for a particular \index{Artefact, Data}Data Artefact, or collection of Data Artefacts.}\\\hline
	{\index{Property!Data|textbf}\index{Data!Property|see{Property, Data}}Data Property} & {A characteristic that can be exhibited by a \index{Artefact, Data}Data Artefact.}\\\hline
	{\index{Safety Assessment!Data|textbf}\index{Data!Safety Assessment|see{Safety Assessment, Data}}Data Safety Assessment} & {The process of explicitly considering data as part of a \index{Safety Assessment!System}system safety assessment, via the means of \index{Artefact, Data}Data Artefacts, \Glspl{Data Property} and Data Safety Assurance Levels.}\\\hline
	{\index{Data Safety Assurance Level|see{Assurance Level, Safety, Data}}\index{Assurance Level!Safety!Data}Data Safety Assurance Level} & {An indication of the level of rigour with which relevant \glspl{Data Property} should be demonstrated for appropriate \index{Artefact, Data}Data Artefacts.}\\\hline
	{\index{Safety Requirement!Data|textbf}\index{Data!Safety Requirement|see{Safety Requirement, Data}}Data Safety Requirement} & {A requirement to implement an approach specifically designed to achieve, maintain or demonstrate a \gls{Data Property} (or Properties) for a given \index{Artefact, Data}Data Artefact (or Artefacts).}\\\hline
    {\index{Response|textbf}Response} & {The way in which an identified risk is addressed; possible responses include avoid/eliminate, treat, or accept as sufficiently low.}\\\hline
	{\index{Stakeholder|textbf}Stakeholder} & {An individual or organization that has some relationship to the system, possibly including a power of veto.}\\\hline
    {\index{Treatment!Risk|textbf}Treatment} & {An action taken to reduce or control risk.}\\\hline
\end{longtable}

A more comprehensive list of definitions, including descriptive definitions, is included at \autoref{bkm:acronyms}.
%-- We don't currently have any normative abbreviations
%\subsection{Abbreviations}
%\textcolor{blue}{XXX TO COMPLETE XXX}
