%================================================================================
%       Safety Critical Systems Club - Data Safety Initiative Working Group
%================================================================================
%                       DDDD    SSSS  IIIII  W   W   GGGG
%                       D   D  S        I    W   W  G   
%                       D   D   SSS     I    W W W  G  GG
%                       D   D      S    I    WW WW  G   G
%                       DDDD   SSSS   IIIII  W   W   GGG
%================================================================================
%               Data Safety Guidance Document - LaTeX Source File
%================================================================================
%
% Description:
%   Objectives and Outputs section.
%
%================================================================================
\section{Objectives and Outputs (Normative)} \label{bkm:objectivesoutputs}

\dsiwgSectionQuote{The goal is to turn data into information, and information into insight.}{Carly Fiorina}

\subsection{Establish Context}
\subsubsection{Objectives}
\begin{enumerate}[label=\color{dsiwgAccentColour}{1-\arabic*}]
	\item System context and intended use SHALL be established.
	\item Key \index{Stakeholder}\glspl{stakeholder} SHALL be identified.
	\item \index{Artefact, Data}\Glspl{data artefact} SHALL be identified.
	\item Interfaces SHALL be defined and managed.
	\item A \index{Safety Assessment!Data}\gls{safety assessment} SHALL be planned.
\end{enumerate}

\subsubsection{Outputs}
\begin{itemize}
	\item A description of the system and its intended use. This SHOULD include an estimate of the level of data-related risks.
	\item A list of key \index{Stakeholder}\glspl{stakeholder} for data safety activities.
	\item A collection of \index{Artefact, Data}\glspl{data artefact}, described at an appropriate level of detail.
	\item An interface control plan or list of control measures. This MAY include a list of \index{Data!Owner}\glspl{data owner}, linked to \index{Artefact, Data}\glspl{data artefact}.
	\item A plan for the remaining parts of the \gls{safety assessment}.
\end{itemize}

\subsection{Identify Risks}
\subsubsection{Objectives}
\begin{enumerate}[label=\color{dsiwgAccentColour}{2-\arabic*}]
	\item Historical data-related accidents and incidents SHALL be reviewed.
	\item Unintended behaviour resulting from data SHALL be identified and analysed.
	\item Risks SHALL be identified and linked to \index{Artefact, Data}\glspl{data artefact} and \index{Property!Data}\glspl{data property}.
\end{enumerate}

\subsubsection{Outputs}
\begin{itemize}
	\item A description of the process used for risk identification.
	\item A list of risks, linked to \index{Artefact, Data}\glspl{data artefact} and \index{Property!Data}\glspl{data property}.
	\item An updated plan for the remaining parts of the \index{Safety Assessment!Data}\gls{safety assessment}, if required.
\end{itemize}

\subsection{Analyse Risks}
\subsubsection{Objectives}
\begin{enumerate}[label=\color{dsiwgAccentColour}{3-\arabic*}]
	\item \Glspl{dsal} SHALL be established.
	\item \Glspl{dsal} SHALL be justified.
	\item \Glspl{dsal} SHALL be incorporated into system safety activities.
\end{enumerate}

\subsubsection{Outputs}
\begin{itemize}
	\item A \gls{dsal} and supporting justification\index{Justification, Safety} for each risk identified in the previous phase.
	\item An updated plan for the remaining parts of the \index{Safety Assessment!Data}\gls{safety assessment}, if required.
\end{itemize}

\subsection{Evaluate and Treat Risks}
\subsubsection{Objectives}
\begin{enumerate}[label=\color{dsiwgAccentColour}{4-\arabic*}]
	\item \index{Safety Requirement!Data}\Glspl{data safety requirement} SHALL be established and elaborated.
	\item Methods used to provide data safety assurance SHALL be defined and implemented.
	\item Compliance with \index{Safety Requirement!Data}\glspl{data safety requirement} SHALL be demonstrated.	
\end{enumerate}

\subsubsection{Outputs}
\begin{itemize}
	\item A record of the agreed \index{Response}\glspl{response} to each of the identified risks, along with supporting justification\index{Justification, Safety}.
	\item A list of \index{Safety Requirement!Data}\glspl{data safety requirement} that follow from these \index{Response}\glspl{response}.
	\item A record of the \index{Treatment!Risk}\gls{treatment} adopted for each of the identified risks, including evidence that the \index{Treatment!Risk}\gls{treatment} has been successfully implemented.
	\item An assessment as to whether the risk has been suitably mitigated\index{Mitigation} (and, if not, plans for further \gls{mitigation}\index{Mitigation} activities).
\end{itemize}
