%================================================================================
%       Safety Critical Systems Club - Data Safety Initiative Working Group
%================================================================================
%                       DDDD    SSSS  IIIII  W   W   GGGG
%                       D   D  S        I    W   W  G   
%                       D   D   SSS     I    W W W  G  GG
%                       D   D      S    I    WW WW  G   G
%                       DDDD   SSSS   IIIII  W   W   GGG
%================================================================================
%               Data Safety Guidance Document - LaTeX Source File
%================================================================================
%
% Description:
%   Data Safety Management Plan section.
%
%================================================================================
\section{\ecapitalisewords{\glsentrylong{dsmp}} (Informative)} \label{bkm:plan}

\dsiwgSectionQuote{Things get done only if the data we gather can inform and inspire those in a position to make a difference.}{Mike Schmoker}

This section gives a suggested \gls{dsmp} table of contents. It is expected that this will be needed only for aspects not already covered in a \gls{smp}, or similar. It can be merged with an \gls{smp}, if appropriate. However it may be useful to consider the distinct data perspective by using a \gls{dsmp} as well as an \gls{smp}. Regardless, a close connection should be maintained between the \gls{smp} and the \gls{dsmp}.

\Gls{dsmp} suggested contents:
\begin{enumerate}
  \item Introduction:
  \begin{itemize}
    \item Scope and Context (Sets the scene, describes the project, scenario, concept of operations, etc.);
    \item Boundaries and Interfaces (Describes the main interfaces and exchanges, with a scope boundary diagram.);
    \item
      Derived requirements (System-level requirements which impact the data safety process);
    \item \index{Data!Owner}Owners (Who owns the data under consideration as it progresses through the system?);
    \item Producers / Consumers (Who are the producers and consumers of the data the system inputs and outputs?);
    \item Assumptions;
    \item References; and
    \item Abbreviations and Acronyms.
  \end{itemize}
  \item Analysis of Assigned \gls{dsal}\index{Assurance Level!Data} and \gls{odr} Level (Implications of the data analyses.):
  \begin{itemize}
    \item \Gls{sil}, etc., Implications (What impact does the \gls{dsal}\index{Assurance Level!Data} have on the required \gls{sil}, or similar measure, and vice versa?);
    \item Development Implications (Are there any special development considerations? Derived from the \gls{sil} if there is one, otherwise what is deemed necessary for this system.);
    \item Verification Implications (Derived from the \gls{sil} if there is one, otherwise what is deemed necessary for this system.);
    \item Assurance Implications (Derived from the \gls{sil} if there is one, otherwise what is deemed necessary for this system.); and
    \item Process / Procedure Implications (Derived from the \gls{sil} if there is one, otherwise what is deemed necessary for this system.).
  \end{itemize}
\item
  Categories
  of Safety Data in Scope (A list of all the categories\index{Category!Data} to be considered in the system context.).
  \item Data Requirements Analysis:
  \begin{itemize}
    \item Lifecycles\index{Lifecycle!Data} (What data lifecycles are to be used?);
    \item Specific Targets (Are there any qualitative or quantitative targets for the data?); and
    \item Security Considerations (How will security be managed in this context? Are there any security / safety conflicts? Are there any security-related causes of data hazards?).
  \end{itemize}
  \item Management Approach (How will the organization manage the data safety risks?):
  \begin{itemize}
    \item Organization;
    \item Responsibilities;
    \item Authorisations; and
    \item Approvals and Signoffs.
  \end{itemize}
  \item Justification Approach (How will the safe usage of the data be justified, e.g., as part of the Safety Case Report?). 
  \item Analyses / Verifications to be Performed (What analyses or checks are to be performed on the data?).
  \item Documents to be Produced (The list of documents to be produced related to data aspects.).
  \item Appendix: \gls{dsal}\index{Assurance Level!Data} Guidelines Response (Tailored version of the tables from this document. What is considered applicable / useful and what is not?).
\end{enumerate}
